Predicting the 2024 US Presidential Election with a Model-Based Forecast

### 2024-10-29 paper/paper.qmd
update introduction

The presidential election has a significant impact on national and global policies, which makes election forecasting a valuable tool for understanding political outcomes. However, individual polls often contain biases and short-term fluctuations that can obscure long-term trends. To deal with these challenges, we use a ‘poll-of-polls’ approach that combines multiple polls to provide more stable predictions of voter preferences over time. In this paper, we use national polling data from the period leading up to the 2024 U.S. presidential election to generate a logistic model that predicts the winner while capturing changes in public opinion throughout the election period.

The model estimates the probability of a candidate winning an election based on aggregated polling data. Logistic regression is used for binary outcomes such as election results, which allows us to model the likelihood of each candidate winning as new data becomes available. By taking into account changes in opinion polls and tracking trends in public opinion over time, this model improves the accuracy of predictions.

Our analysis identifies key trends among the major candidates and tracks changes in public sentiment throughout the campaign. The results suggest that aggregated polls can reduce some uncertainty by smoothing short-term fluctuations between surveys, but uncertainty cannot be eliminated. A sudden change in public opinion can still change the path of an election campaign, which highlights the importance of continuous polling monitoring.

A reliable election forecast can influence public expectations, and campaign strategies, and increase the visibility of reporting on election trends. An accurate forecast can also increase voter participation by providing individuals with a clearer perspective of the electoral landscape, and help them participate more effectively in the process of democracy. Prediction models like ours not only help with resource allocation for campaigns but also support the wider public by providing them with a clearer understanding of political dynamics.

The structure of the paper is organized as follows: following @sec-intro, @sec-data presents the data collection and cleaning process, along with an overview of the variables used in the analysis. @sec-model introduces the forecasting models, explaining why the selected models are suitable for predicting election outcomes based on aggregated polling data. Then @sec-result presents the main findings, including detailed crime trends for each neighborhood and year. Finally, @sec-discussion provides the results, highlighting key trends and predictions. Eventuallt, @sec-discussion concludes with a discussion of the findings, evaluating the reliability of the forecasts and identifying potential limitations of the models.

>>>>

The results of presidential elections influence national and international policies and determine governance and economic priorities. An accurate election forecast is a valuable tool for political strategists, the media, and the public. The 2024 U.S. presidential election is expected to be highly competitive, with Donald Trump and Kamala Harris as the leading candidates. In this paper, we use national polling data from the period leading up to the 2024 U.S. presidential election to generate a linear regression model that predicts the winner while capturing changes in public opinion throughout the election period.

This study estimates the probability of each candidate winning based on aggregate polling data. By combining the results, we reduce the bias and short-term volatility of individual polls. By using a linear model, we estimate the probability of a Trump victory to be between 45% and 55%, which depends on state-by-state variations.

Our findings suggest that while Trump has a competitive chance, Harris is projected to win key states and receive a majority of electoral votes. Harris will receive 239 electoral votes to Trump's 98 votes. This is important because accurate election forecasts help reduce uncertainty, which allows political stakeholders to allocate resources efficiently and gives the public a better understanding of likely outcomes. Aggregate polling reduces uncertainty and provides a clearer picture of potential outcomes.

The structure of the paper is organized as follows: following @sec-intro, @sec-data presents the data collection and cleaning process, along with an overview of the variables used in the analysis. @sec-model introduces the forecasting models, explaining why the selected models are suitable for predicting election outcomes based on aggregated polling data. Then @sec-result presents the main findings, including detailed crime trends for each neighborhood and year. Finally, @sec-discussion provides the results, highlighting key trends and predictions. Eventually, @sec-discussion concludes with a discussion of the findings, evaluating the reliability of the forecasts and identifying potential limitations of the models.
